The algebraic hierarchy in type theory using type classes

### 2010-06-22 jar.tex

With these instances in place, we can now define the improved natural numbers specification:
\begin{lstlisting}
  Class Naturals A {e plus mult zero one} `{!NaturalsToSemiRing A} :=
    { naturals_ring:> @SemiRing A e plus mult zero one
    ; naturals_to_semiring_mor:> Π `{SemiRing B}, SemiRing_Morphism (naturals_to_semiring A B)
    ; naturals_initial:> Initial (bundle_semiring A) }.
\end{lstlisting}
Basing theory and programs on this abstract interface instead of on specific implementation (such as the ubiquitous Peano naturals \lstinline|nat| in the Coq standard library) is not only cleaner mathematically, but also facilitates easy swapping between implementations. And this benefit is far from theoretical, as diverse and non-contrived representations of the natural numbers are abound (think unary, binary, factor multisets, arrays of native machine words, etc.).

>>>>

With these instances in place, we can now define the improved natural numbers specification:
\begin{lstlisting}
  Class Naturals A {e plus mult zero one} `{!NaturalsToSemiRing A} :=
    { naturals_ring:> @SemiRing A e plus mult zero one
    ; naturals_to_semiring_mor:> Π `{SemiRing B}, SemiRing_Morphism (naturals_to_semiring A B)
    ; naturals_initial:> Initial (bundle_semiring A) }.
\end{lstlisting}
Basing theory and programs on this abstract interface instead of on specific implementation (such as the ubiquitous Peano naturals \lstinline|nat| in the Coq standard library) is not only cleaner mathematically, but also facilitates easy swapping between implementations. And this benefit is far from theoretical, as diverse and non-contrived representations of the natural numbers are abound (unary, binary, factor multisets, arrays of native machine words, etc.).
